---
title: "WUP_04_Alt"
author: "Tyler Hamblin"
date: "Fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and the RACE of the person taking the test   
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two different RACES

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

This string of code allows the character variables to be changed into factor variables under the section "mydata". This type of question requires a numerical variable (WORDSUM) to be explained by a three level factor variable (RACE). A group of helper files will allow us to complete this analysis of numerical and factorial variables. 


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=RACE,y=WORDSUM))

basicCN + geom_boxplot()
```

Simply looking at this plot doesn't allow us to narrow in to where the statistical difference is because all three of the boxes overlap at some point. Although we can assume that there is some variation in data, the question asks us where that occurs so we must run more tests to accurately and specifically answer. 

There is not enough variation between OTHER and WHITE, BLACK and OTHER, or especially BLACK and WHITE because the boxes and medians (black lines) are overlapping or aligned. 


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(WORDSUM~RACE,data=mydata)
```

According to the numbers that printed out we can see that there is some distinction at the quartile marks especially in the "OTHER" category that seems to perform the lowest every time. With the lowest mean, maximum, and median we keep this in mind for the next test to see if this statistical suggestion holds true.



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~RACE,data=mydata)
summary.aov(model1)
```

This one way anova test collects our p value for us, 0.0182, which confirms that we need to continue running more tests. 


```{r}
model1 <- aov(WORDSUM~RACE,data=mydata)
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

p = 0.0182

This p value suggests a significant result because of it being so minute but needs to be further investigated with a multiple comparisons test to decide whether or not your race has to do with your ability to perform on this test. 

This final version of the plot shows us that at the value of zero there is no difference because White-Other + White-Black and White-Black + Black-Other  zero is within the confidence interval. White-Other and Black-Other appear to be the only two categories that don't overlap and therefore do have a noticeable distinction.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

$H_0$ There is no difference between the average score on the WORDSUM test and the RACE of the person taking the test   
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two different RACES

The $H_0$ (null hypothesis) can be rejected based on our results at the end of each graph that suggest a continued statistical variation in the data. We in turn accept the $H_a$ (alternate hypothesis) which states that at least two different races receive different results on their average score. We can especially see this on the confidence level test in which the Black-Other and White-Other limits do not extend past each other. Along with the original box plot, a favstats command, and a t-test to set up the final graph, all of which the helper file 'Numerical~Cat' provided for us, we can be confident in this conclusion. 

